---
format:
  dashboard:
    logo: images/Paris2024_Mono_Blanc.png
    orientation: rows
    main-font: gowun-dodum
    expandable: false

execute:
  echo: false
  warning: false
  error: false
---

<!-- H1 = PAGE -->

# Olympics in numbers

```{r}
#| label: r-session-setup

# data wrangling
library(tidyverse)
library()
# html and icons
library(htmltools)
library(bslib)
library(bsicons)
library(icons)
# tables
library(gt)
#library(tabulapdf) # from pdf
# maps + country codes
library(rnaturalearth)
library(sf)
library(cartogram)
library(countrycode)
library(maps)
library(geosphere)
#library(tidygeocoder)
# plots
library(ggiraph)
library(scatterpie)
library(gganimate)
library(ggtext)
library(scales)
library(ggarrow)
```

```{r}
#| label: general-visuals

### Medal colours
col_medals <- c("#FCC65B", "#E5E5E5", "#E0BD95")
names(col_medals) <- c("Gold", "Silver", "Bronze")

### Gender colours
col_genders <- c("#304BBC", "#A53771")

### Paris 2024 colours
col_paris <- c("#f292b7", "#d6be76", "#1d87c2", "#002e54", "#a499c3")
```


```{r}
#| label: get-data

### All NOCs
all_noc <- read.delim("data/nocs.csv", sep = ",", header = T)
# These include historical territories as well - exclude them
noc_present <- filter(all_noc, note == "P")

### Participating athletes
athletes_dat <- read.delim("data/athletes.csv", sep = ",", header = T)

### Nations with medals
nations_dat <- read.delim("data/medals_total.csv", sep = ",", header = T)

### Medalists
medalists_dat <- read.delim("data/medallists.csv", sep = ",", header = T)
# this will unify terms used for the medal types
medalists_dat$medal_type <- gsub(" Medal", "", medalists_dat$medal_type)

### Schedule
schedule_dat <- read.delim("data/schedules.csv", sep = ",", header = T)

### Preprocessed data
load("data/female_historical.RData")
load("data/torch_route_coordinates.RData")
```

```{r}
#| label: overview-stats

### AIM: Get NOCs vs qualified vs with at least one medal

# All NOCs
n_noc_all <- nrow(noc_present)

# NOCs qualified
n_noc_qualified <- length(unique(athletes_dat$country_code))

# NOCs with medals
n_noc_medals <- length(unique(nations_dat$country_code))


### AIM: Medal-winning competitions

# Total of medal sets awarded
n_awards <- sum(schedule_dat$event_medal)
```

## ROW: Intro {height="20%"}

::: {.div-intro style="font-size: 0.8em;"}
The 33^rd^ Summer Olympic Games - or simply **Paris 2024** - came back to the capital of France after the 1900 and 1924 Games.

Developments of the past century were reflected in the focus of the Games' organizers: gender equality and sustainability. The Paris Games also stood out with their Opening Ceremony - for the first time in modern Olympic history, it was held outside of a stadium. River Seine became the venue.

Olympics are a global event, raising a lot of interest but also controversies. This dashboard celebrates sports achievments of the participants.
:::

<!-- And torch and venues???-->

## ROW: Subset of cards

### COL : Card 1 {height="20%"}
```{r}
#| label: valuebox-medal-competitions

bslib::value_box(
  title = tags$h3("Medal ceremonies", style = "cont-size: 20px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("award-fill"),
  value = tags$p(n_awards, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Card 2 {height="20%"}

```{r}
#| label: valuebox-all-nocs

bslib::value_box(
  title = tags$h3("Registered NOCs", style = "cont-size: 20px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("globe2"),
  value = tags$p(n_noc_all, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Card 3 {height="20%"}

```{r}
#| label: valuebox-nocs-qual

bslib::value_box(
  title = tags$h3("Qualified NOCs", style = "cont-size: 20px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("check2-all"),
  value = tags$p(n_noc_qualified, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Card 4 {height="20%"}

```{r}
#| label: valuebox-nocs-medals

bslib::value_box(
  title = tags$h3("NOCs with medalists", style = "cont-size: 20px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("award"),
  value = tags$p(n_noc_medals, style = "font-size: 42px; font-face: bold;")
)
```

## ROW: Genders

```{r}
#| label: get-athletes-by-gender

n_men <- athletes_dat |> 
  filter(gender == "Male" & function. == "Athlete") |> 
  select(country, name, disciplines) |> 
  distinct() |> 
  nrow()

n_women <- athletes_dat |> 
  filter(gender == "Female" & function. == "Athlete") |> 
  select(country, name, disciplines) |> 
  distinct() |> 
  nrow()
```

### COL: Men's card {height="20%"}

```{r}
#| label: valuebox-male-athletes

bslib::value_box(
  title = tags$h3("Male athletes", style = "cont-size: 20px;"),
  theme = "#000",
  showcase = bs_icon("gender-male"),
  value = tags$p(n_men, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Women's card {height="20%"}

```{r}
#| label: valuebox-female-athletes

bslib::value_box(
  title = tags$h3("Female athletes", style = "cont-size: 20px;"),
  theme = "bg-theme-paris-pink",
  showcase = bs_icon("gender-female"),
  value = tags$p(n_women, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Facts about the Games {height="50%"}



- mascot
Named Olympic Phryge, the Paris 2024 mascot is based on the traditional small Phrygian hats that the mascots are shaped after.

Its name and design were chosen as symbols of freedom and to represent allegorical figures of the French Republic.


The Paris Games were exceptional in many other aspects - let's explore a selection together. 

- torch
Carried by 10,000 people over thousands of miles, the torch, crafted from metal by steel giant Arcelor Mittal, resembles an oval iron pipe with a rounded center.

Upon its arrival, the torch will have been carried to more than 400 locations, including France's overseas territories, until the opening of the event on July 26.

## ROW: Plot genders in time {height="50%" width="50%"}

```{r}
#| label: figure-historical-female-participation

# n of frames ~ n of years with data
n_frames = length(unique(female_dat$Year))

gender_plot_title <- paste0("**<span style='color:", col_paris[3], "'>Male</span> and ", "<span style='color:", col_paris[1], "'>female</span> athletes in Summer Olympics: 1900-2024**")

# long format for plotting
p_genders <- female_dat |> 
  pivot_longer(cols = c("Prcnt_women_participants", "Prcnt_men_participants"), values_to = "Prcnt", names_to = "Gender") |> 
  ggplot(aes(x = Year, y = Prcnt, group = Gender)) +
  geom_line(aes(colour = Gender), linewidth = 1.25) +
  scale_colour_manual(values = c("Prcnt_men_participants" = col_paris[3], "Prcnt_women_participants" = col_paris[1])) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = unique(female_dat$Year)) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = gender_plot_title) +
  theme_minimal() +
  theme(plot.title = element_markdown(hjust = 0.5),
        axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        aspect.ratio = 0.5,
        legend.position = "none")

anim_genders <- p_genders +
  transition_reveal(along = Year)

animate(anim_genders, nframes = n_frames, duration = 7)
```

<!-- -------------------------------------------------------------------------- -->

<!-- -------- -->
<!-- New page -->

# Participating nations {orientation="columns"}

{{< include _TopNations.qmd >}}

# Top medallists {orientation="columns"}

{{< include _TopMedallists.qmd >}}


# Coaches

# Torch relay

{{< include _TorchRoute.qmd >}} {orientation="columns"}


<!-- IDEAS: Total men and women, disciplines with most events/medals? -->

# About

**Motivation:**  

I made this website to practice use of Quarto dashboards and deployment to public.

<!-- elective RData/qs or pdfs??? -->