---
format:
  dashboard:
    logo: images/Paris2024_Mono_Blanc.png
    orientation: rows
    main-font: gowun-dodum
    expandable: false

execute:
  echo: false
  warning: false
  error: false
---

<!-- H1 = PAGE -->

# Olympics in numbers

```{r}
#| label: r-session-setup

# data wrangling
library(tidyverse)
library()
# html and icons
library(htmltools)
library(bslib)
library(bsicons)
library(icons)
# tables
library(gt)
#library(tabulapdf) # from pdf
# maps + country codes
library(rnaturalearth)
library(sf)
library(cartogram)
library(countrycode)
library(maps)
library(geosphere)
#library(tidygeocoder) # get geographical coordinates for countries/places/addresses
# plots
library(ggiraph)
library(scatterpie)
library(gganimate)
library(ggrepel)
library(ggtext)
library(scales)
library(arcdiagram)
library(igraph)
library(ggraph)
library(ggalluvial)
library(RColorBrewer)
library(gridExtra)
library(gtable)
```

```{r}
#| label: general-visuals

### Medal colours
col_medals <- c("#FCC65B", "#E5E5E5", "#E0BD95")
names(col_medals) <- c("Gold", "Silver", "Bronze")

### Gender colours
col_genders <- c("#304BBC", "#A53771")

### Paris 2024 colours
col_paris <- c("#f292b7", "#d6be76", "#1d87c2", "#002e54", "#a499c3")
```


```{r}
#| label: get-data

### All NOCs
all_noc <- read.delim("data/nocs.csv", sep = ",", header = T)
# These include historical territories as well - exclude them
noc_present <- filter(all_noc, note == "P")

### Participating athletes
athletes_dat <- read.delim("data/athletes.csv", sep = ",", header = T)

### Nations with medals
nations_dat <- read.delim("data/medals_total.csv", sep = ",", header = T)

### Medalists
medalists_dat <- read.delim("data/medallists.csv", sep = ",", header = T)
# this will unify terms used for the medal types
medalists_dat$medal_type <- gsub(" Medal", "", medalists_dat$medal_type)

### Schedule
schedule_dat <- read.delim("data/schedules.csv", sep = ",", header = T)

### Preprocessed data
load("data/female_historical.RData")
load("data/torch_route_coordinates.RData")
```

```{r}
#| label: overview-stats

### AIM: Get NOCs vs qualified vs with at least one medal

# All NOCs
n_noc_all <- nrow(noc_present)

# NOCs qualified
n_noc_qualified <- length(unique(athletes_dat$country_code))

# NOCs with medals
n_noc_medals <- length(unique(nations_dat$country_code))


### AIM: Medal-winning competitions

# Total of medal sets awarded
n_awards <- sum(schedule_dat$event_medal)
```

## ROW: Intro {height="21%"}

::: {#div-intro style="font-size: 0.8em;"}
The 33^rd^ Summer Olympic Games - or simply **Paris 2024** - came back to the capital of France after the 1900 and 1924 Games.

Developments of the past century were reflected in the focus of the Games' organizers: gender equality and sustainability. The Paris Games also stood out with their Opening Ceremony - for the first time in modern Olympic history, the ceremony was held outside of a stadium. River Seine became the venue.

Olympics are a global event, raising a lot of interest but also some controversies. This dashboard deliberately focus on sport achievements only.
:::


## ROW: Subset of cards

### COL : Card 1
```{r}
#| label: valuebox-medal-competitions

bslib::value_box(
  title = tags$h3("Medal ceremonies", style = "font-size: 18px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("award-fill"),
  value = tags$p(n_awards, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Card 2

```{r}
#| label: valuebox-nocs-qual

bslib::value_box(
  title = tags$h3("Qualified NOCs", style = "font-size: 18px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("check2-all"),
  value = tags$p(n_noc_qualified, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Card 3

```{r}
#| label: valuebox-all-nocs

bslib::value_box(
  title = tags$h3("Registered NOCs", style = "font-size: 18px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("globe2"),
  value = tags$p(n_noc_all, style = "font-size: 42px; font-face: bold;")
)
```


### COL: Card 4

```{r}
#| label: valuebox-nocs-medals

bslib::value_box(
  title = tags$h3("Medal-winning NOCs", style = "font-size: 18px;"),
  theme = "bg-theme-paris-dblue",
  showcase = bs_icon("award"),
  value = tags$p(n_noc_medals, style = "font-size: 42px; font-face: bold;")
)
```

## ROW: Explanation {height="6%"}

:::{#div-noc style="text-align: right; font-size: 0.8em;"}
*NOC: national olympic committee*
:::

## ROW: Genders

```{r}
#| label: get-athletes-by-gender

n_men <- athletes_dat |> 
  filter(gender == "Male") |> 
  select(country, name, disciplines) |> 
  distinct() |> 
  nrow()

n_women <- athletes_dat |> 
  filter(gender == "Female") |> 
  select(country, name, disciplines) |> 
  distinct() |> 
  nrow()
```

### COL: Men's card

```{r}
#| label: valuebox-male-athletes

bslib::value_box(
  title = tags$h3("Male athletes", style = "font-size: 18px;"),
  theme = "#000",
  showcase = bs_icon("gender-male"),
  value = tags$p(n_men, style = "font-size: 42px; font-face: bold;")
)
```

### COL: Women's card

```{r}
#| label: valuebox-female-athletes

bslib::value_box(
  title = tags$h3("Female athletes", style = "font-size: 18px;"),
  theme = "bg-theme-paris-pink",
  showcase = bs_icon("gender-female"),
  value = tags$p(n_women, style = "font-size: 42px; font-face: bold;")
)
```

## ROW: Plot genders in time {height="45%"}

```{r}
#| label: figure-historical-female-participation

# n of frames ~ n of years with data
n_frames = length(unique(female_dat$Year))

gender_plot_title <- paste0("**<span style='color:", col_paris[3], "'>Male</span> and ", "<span style='color:", col_paris[1], "'>female</span> athletes in Summer Olympics: 1900-2024**")

# long format for plotting
p_genders <- female_dat |> 
  pivot_longer(cols = c("Prcnt_women_participants", "Prcnt_men_participants"), values_to = "Prcnt", names_to = "Gender") |> 
  ggplot(aes(x = Year, y = Prcnt, group = Gender)) +
  geom_line(aes(colour = Gender), linewidth = 1.25) +
  scale_colour_manual(values = c("Prcnt_men_participants" = col_paris[3], "Prcnt_women_participants" = col_paris[1])) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = unique(female_dat$Year)) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = gender_plot_title) +
  theme_minimal() +
  theme(plot.title = element_markdown(size = 15, hjust = 0.5),
        axis.text.x = element_text(size = 14, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(size = 14),
        axis.title = element_blank(),
        aspect.ratio = 0.5,
        legend.position = "none")

anim_genders <- p_genders +
  transition_reveal(along = Year)

animate(anim_genders, nframes = n_frames, duration = 14, end_pause = 7)
```

<!-- -------------------------------------------------------------------------- -->

<!-- -------- -->
<!-- New page -->

# Participating nations {orientation="columns"}

{{< include _TopNations.qmd >}}


# Athletes & coaches {orientation="columns"}

{{< include _Coaches.qmd >}}


# Top medallists {orientation="columns"}

{{< include _TopMedallists.qmd >}}


# Torch relay {orientation="rows"}

{{< include _TorchRoute.qmd >}}


<!-- --------------------------------------------------------- -->

# About

**Motivation:**  

I started this project to practice use of Quarto dashboards and deployment to public. Besides some of my favourite packages, I aimed to use packages and visualizations that I did not have a chance to use for my work yet.

## Acknowledgments

**Data source:**  
- Olympics statistics: https://www.kaggle.com/datasets/piterfm/paris-2024-olympic-summer-games?resource=download  
- HEX colour codes from [Olympics Paris 2024 Color Palette](https://www.color-hex.com/color-palette/1049914)

**Images:**  
- https://press.paris2024.org/photos/paris2024-olyembleme-rvb-mono-blanc-2021-png-a812-7578a.html

**Inspirations:**  
The Internet is full of useful and inspiring resources which I probably used without being truly aware. Nonetheless, there are video tutorials and inspirational people I would like to mention:  
- A [dashborad tutorial](https://www.youtube.com/watch?v=27Njh1i9-oM  ) by Melissa Van Bussel (ggnot2)  
- A [tutorial series](https://www.youtube.com/watch?v=HW7QbqI4fH0  ) by Mine Çetinkaya-Rundel (Posit)  
- tips and tutorials by Yan Holtz, Albert Rapp and Isabella Velásquez  

